An open reproducible framework for the study of the iterated prisoner's dilemma

### 2016-02-09 tex/paper.tex
Various minor tweaks after a read through.

\begin{abstract}
\noindent
    The Axelrod Python library is an open source package that allows for
    reproducible game theoretic research in to the iterated Prisoner's Dilemma.
    This area of research began in the 1980 but suffers from a lack of
    documented and tested code. This library aims to address this.\\

\noindent
    With a growing collection of 120 strategies the library is also a platform
    for an original tournament that in itself is of interest to the game
    theoretic community.
\end{abstract}

@@

Several iterated prisoner's dilemma tournaments have generated much interest,
including Axelrod's original tournaments, two 2004 anniversary tournaments, and
the Stewart and Plotkin 2012 tournament \cite{Stewart2012} following the
discovery of zero-determinant strategies.  Subsequent research has spawned an
enormous number of papers, but rarely are the results reproducible. Among
well-known tournaments, in only one case is the full original source code
available (Axelrod's second tournament, in FORTRAN); in no cases is the
available code well-documented, easily modifiable, or released with significant
test suites.

@@

\begin{itemize}
    \item Open: all code is released under an MIT license;
    \item Reproducible and well-tested: at the time of writing there is an excellent level of
        integrated tests with 99.59\% coverage
    \item Well-documented: all features of the library are documented for ease of
        use and modification
    \item Extensive: over 100 strategies are included, with infinitely-many
    strategies available in the case of parameterized strategies
    \item Extensible: easy to modify to include new strategies and to run new tournaments
\end{itemize}

Before describing the package in more detail in
Section~\ref{sec:description-of-axelrod-python}, an overview of some previous
Iterated Prisoner's Dilemma research will be given in
Section~\ref{sec:review}.

@@

\begin{itemize}
    \item If both choose \(C\), they receive a payoff of \(R\)
        (\textbf{R}eward);
    \item If both choose \(D\), they receive a payoff of \(P\)
        (\textbf{P}punishment);
    \item If one chooses \(C\) and the other \(D\), the defector receives a
        payoff of \(T\) (\textbf{T}emptation) and the cooperator a payoff of
        \(S\) (\textbf{S}ucker).
\end{itemize}

@@

The game of (\ref{equ:one-shot}) is called the Prisoner's Dilemma. Numerical
values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature. Axelrod's
tournaments (and further implementations of these) are sometimes referred to as
Iterated Prisoner's Dilemma (IPD) tournaments. An incomplete overview of
published tournaments is given in Table~\ref{tab:tournaments}.

@@

Furthermore, \cite{Isaac2008} Python is described as an appropriate language for the
reproduction of Iterated Prisoner's dilemma tournaments due to its object
oriented nature and readability.

The library itself is available at
\url{https://github.com/Axelrod-Python/Axelrod}. This is a hosted git
repository. Git is a popular version control system which is one of the
recommended aspects of reproducible research \cite{Crick2014a, Sandve2013}.

@@

Figure~\ref{fig:demo_tournament_commands} just shows the very basic utilisation
of the library and further details can be found at the online documentation:
\url{http://axelrod.readthedocs.org}.

As stated in Section~\ref{sec:introduction} one of the main goals of the library
is to allow for the easy contribution of strategies. To do this requires the
writing of a simple Python class (which can inherit from other predefined
classes). Full contribution guidelines can be found in the documentation.
As an example, Figures~\ref{fig:grudger} and~\ref{fig:grudger_test} show the source code for
the Grudger strategy as well as its corresponding test.

@@

\begin{figure}[!hbtp]
	\centering
	\includegraphics[width=.9\textwidth]{../img/tournament.pdf}
	\caption{Results from the library tournament (2015-12-19)}
	\label{fig:tournament}
\end{figure}

@@

The above example is an incomplete illustration of the mapping for \(m=8,
n=2\). Intuitively, this state space uses the initial plays of the opponent to
gain some information about its intentions whilst still taking in to account
the recent play. The actual winning strategy is an instance of the framework
for \(m=n=2\) for which an evolutionary algorithm has been used to train it.
More details of this can be found in \cite{Jones2015}. In \cite{Knight2015}
experiments are described that evaluate how this particular strategy behaves in
environments other than those in which it was trained and it continues to
perform strongly.

@@

\begin{itemize}
    \item A closer look at zero determinant strategies, showing that
        extortionate strategies obtain a large number of wins \textit{but} do
        not perform particularly well. This is relevant given the findings of
        \cite{Stewart2012} in which zero determinant strategies are shown to be
        able to perform better than any other strategy. This finding extends
        to noisy tournaments (which are also implemented in the library).
    \item This negative relationship between wins and performance does not
        generalise. There are some strategies that perform well, both in terms
        of matches won and overall performance: Back stabber, Double crosser,
        Looker Up, and Fool Me Once. These strategies continue to perform well in noisy
        tournaments, however some of these have knowledge of the length of the
        game (Back stabber and Double crosser). This is not necessary to rank
        well in both wins and score as demonstrated by Looker Up and Fool Me
        Once.
    \item Strategies like LookerUp and MetaHunter seem to be generally
        cooperative yet still exploit naive strategies. The MetaHunter strategy
        is a particular type of Meta strategy which uses a variety of other
        strategy behaviours to choose a best action. These strategies perform
        very well in general and continue to do so in noisy tournaments.
\end{itemize}

>>>>

\begin{abstract}
\noindent
    The Axelrod Python library is an open source package that allows for
    reproducible game theoretic research in to the iterated Prisoner's Dilemma.
    This area of research began in the 1980 but suffers from a lack of
    documented and tested code.

\noindent
    With a growing collection of 120 strategies the library is also a platform
    for an original tournament that in itself is of interest to the game
    theoretic community.

\noindent
    This paper describes the iterated Prisoner's Dilemma, the library and
    insights gained from some novel research.
\end{abstract}

@@

Several iterated prisoner's dilemma tournaments have generated much interest,
including Axelrod's original tournaments, two 2004 anniversary tournaments, and
the Stewart and Plotkin 2012 tournament \cite{Stewart2012} following the
discovery of zero-determinant strategies.  Subsequent research has spawned an
enormous number of papers, but rarely are the results reproducible. Among
well-known tournaments, in only one case is the full original source code
available (Axelrod's second tournament \cite{Axelrod1980b}, in FORTRAN); in no
cases is the available code well-documented, easily modifiable, or released with
significant test suites.

@@

\begin{itemize}
    \item Open: all code is released under an MIT license;
    \item Reproducible and well-tested: at the time of writing there is an excellent level of
        integrated tests with 99.37\% coverage
    \item Well-documented: all features of the library are documented for ease of
        use and modification
    \item Extensive: 120 strategies are included, with infinitely-many
    strategies available in the case of parameterized strategies
    \item Extensible: easy to modify to include new strategies and to run new tournaments
\end{itemize}

Before describing the package in more detail in
Section~\ref{sec:description-of-axelrod-python}, an overview of some previous
Iterated Prisoner's Dilemma research will be given in
Section~\ref{sec:review}. Section~\ref{sec:new-strategies-and-implications} will
describe some early research and a conclusion is offered in
Section~\ref{sec:conclusion}.

@@

\begin{itemize}
    \item If both choose \(C\), they receive a payoff of \(R\)
        (\textbf{R}eward);
    \item If both choose \(D\), they receive a payoff of \(P\)
        (\textbf{P}unishment);
    \item If one chooses \(C\) and the other \(D\), the defector receives a
        payoff of \(T\) (\textbf{T}emptation) and the cooperator a payoff of
        \(S\) (\textbf{S}ucker).
\end{itemize}

@@

The game of (\ref{equ:one-shot}) is called the Prisoner's Dilemma. Specific
numerical values of \((R,S,T,P)=(3,0,5,1)\) are often used in the literature.
Axelrod's tournaments (and further implementations of these) are sometimes
referred to as Iterated Prisoner's Dilemma (IPD) tournaments. An incomplete
overview of published tournaments is given in Table~\ref{tab:tournaments}.

@@

Furthermore, in \cite{Isaac2008} Python is described as an appropriate language for the
reproduction of Iterated Prisoner's dilemma tournaments due to its object
oriented nature and readability.

The library itself is available at
\url{https://github.com/Axelrod-Python/Axelrod}. This is a hosted git
repository. Git is a version control system which is one of the
recommended aspects of reproducible research \cite{Crick2014a, Sandve2013}.

@@

Figure~\ref{fig:demo_tournament_commands} shows the very basic utilisation
of the library and further details can be found at the online documentation:
\url{http://axelrod.readthedocs.org}.

As stated in Section~\ref{sec:introduction} one of the main goals of the library
is to allow for the easy contribution of strategies. Doing this requires the
writing of a simple Python class (which can inherit from other predefined
classes). Full contribution guidelines can be found in the documentation.
As an example, Figures~\ref{fig:grudger} and~\ref{fig:grudger_test} show the source code for
the Grudger strategy as well as its corresponding test.

@@

\begin{figure}[!hbtp]
	\centering
	\includegraphics[width=.9\textwidth]{../img/tournament.pdf}
	\caption{Results from the library tournament (2016-02-10)}
	\label{fig:tournament}
\end{figure}

@@

The example of (\ref{equ:example_of_lookerup}) is an incomplete illustration of
the mapping for \(m=8, n=2\). Intuitively, this state space uses the initial
plays of the opponent to gain some information about its intentions whilst still
taking in to account the recent play. The actual winning strategy is an instance
of the framework for \(m=n=2\) for which an evolutionary algorithm has been used
to train it.  More details of this can be found in \cite{Jones2015}. In
\cite{Knight2015} experiments are described that evaluate how this particular
strategy behaves in environments other than those in which it was trained and it
continues to perform strongly.

@@

\begin{itemize}
    \item A closer look at zero determinant strategies, showing that
        extortionate strategies obtain a large number of wins \textit{but} do
        not perform particularly well. This is relevant given the findings of
        \cite{Stewart2012} in which zero determinant strategies are shown to be
        able to perform better than any other strategy. This finding extends
        to noisy tournaments (which are also implemented in the library).
    \item This negative relationship between wins and performance does not
        generalise. There are some strategies that perform well, both in terms
        of matches won and overall performance: Back stabber, Double crosser,
        Looker Up, and Fool Me Once. These strategies continue to perform well in noisy
        tournaments, however some of these have knowledge of the length of the
        game (Back stabber and Double crosser). This is not necessary to rank
        well in both wins and score as demonstrated by Looker Up and Fool Me
        Once.
    \item Strategies like Looker Up and Meta Hunter seem to be generally
        cooperative yet still exploit naive strategies. The MetaHunter strategy
        is a particular type of Meta strategy which uses a variety of other
        strategy behaviours to choose a best action. These strategies perform
        very well in general and continue to do so in noisy tournaments.
\end{itemize}
